The importance of measurement uncertainty in ecological management

### 2018-08-17 manuscripts/manuscript.Rmd
nearly there

```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,  dev="cairo_pdf", cache = FALSE)
library(tidyverse)
library(gridExtra)
```

@@

Imperfect information is ubiquitous in ecological management and conservation decision making. While the pressing concerns of global change have put the spotlight on forecasting, i.e. the uncertainty of the future [e.g. @Petchy2015], management decisions must also contend with uncertainty of the present: How many fish are in the sea today? What regions harbor the most fragile biodiversity? How do we make use of a rapidly expanding but opportunistically collected data, such as in citizen science efforts, which may involve strong sampling biases? Much of our focus in dealing with uncertainty has been through the lens of statistical approaches to model fitting.  The rise of approaches such as Hierarchical Bayesian Modeling have given us sophisticated tools to reflect uncertainty in our data and uncertainty in our knowledge of mechanisms and parameters in our models [].  But estimating a model is not the same as making a decision, particularly when a model leaves us with considerable uncertainty about current and future states. A model can only provide the rules of the chess board.  It is the role of decision theory to provide us with a strategy.

Decision theory has a long history in the ecological literature [e.g @Schaefer1954; @Walters1978; @Clark1990; @Shea1999; @Polasky2011] which has sometimes been overshadowed by recent emphasis on model fitting.  While in principle a model will be updated throughout the decision process (as we will discuss), it is convenient to think of the model as "given," estimated, with uncertainty, using the best available methods. The goal of a decision theory is not maximize some abstract notion of model fit (least squares, likelihood, etc), but a more practical notion of utility (happiness, economic value, or any combination of stakeholder objectives, see @Halpern2013).   Approaches for decision making under uncertainty in ecological systems can be divided into two camps: (1) those based upon "optimal control" solutions, and usually favored by natural resource economists, and (2) those based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011]. While researchers have recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknowledged by heuristic methods, computational barriers to doing so have stymied this progress.  We illustrate how the limitations of these approaches have manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigor of optimization to resolve a long-standing paradox and suggest a more robust approach to management.

@@

In this paper, we examine three strategies to managing a natural population, such as a fishery, under such uncertainty.  This problem has a long history in both theory and practice, where seemingly innoccous simplifying assumptions have led to a deep paradox in how managers should respond to uncertainty.  The precautionary principle [@] would seem to suggest that the greater the uncertainty in our models, the more cautious we should be in management policies.  Yet a well-known mathematical proof by @Reed1979 established that under quite general conditions, the optimal strategy under uncertainty is no different than under no uncertainty at all.  Attempts to resolve this have since only deepened this paradox, arguing that adding additional sources of uncertainty, such as imperfect measurements, should lead to optimal harvest that increases with greater uncertainty [e.g. @Clark1986; @Sethi2005].  Meanwhile, much of actual practice has (perhaps fortunately, as we shall see) ignored these results in favor of a more heuristic application of the precautionary principle.  Here, we show that this paradox can be resolved by uncovering previous shortcuts and tackling the optimal management problem head-on, which we can solve thanks to the computational efficiency of sophisticated algorithms developed more recently in the field of robotics.  This approach allows us to reject this paradox largely in favor of the precautionary principle. Further, our comparisons show that widely recognized existing approaches in both theory and practice are not by themselves sufficient to manage a population under substantial uncertainty over the long term, but our recently borrowed approach from robotics literature can under identical conditions sustain both higher biomass and higher economic yields than those approaches.  This conclusion has important implications not only for fisheries but eocsystem more broadly, underscoring the danger of not correctly accounting for measurement uncertainty when making decisions, and further to the significant opportunity to learn from fields such as robotics when faced with complex and uncertain decisions.

@@

There is one further subtlety arising in this seemingly simple problem statement that we must address. Given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?  Following convention in the optimal control literature (e.g. [@Reed1979; @Clark1990; @Sethi2005]), we will assume the latter; measurement occurs immediately before harvest, followed by recruitment. This assumption has occasionally been reversed to introduce some additional uncertainty from stochastic recruitment immediately prior to setting the harvest quota [e.g. @Clark1986; @Weitzmann2001], which appears to mimic uncertainty in measurements but does not allow uncertainty to accumulate since the measurement step still occurs without error.

@@

@@

We consider the average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.  Each suite of simulations is considered under three sequentially higher measurement error regimes, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated, while stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.  In the absence of either environmental noise or measurement error POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level (see Supplemental material, Figure S1).  The introduction of stochastic growth has a significant negative impact on the TAC strategy, with a mean biomass significantly lower than $B_{MSY}$ (first panel, Figure \ref{sims}; this impact is even more severe for MSY, see Figure S2).  Without measurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.  As measurement error increases in the subsequent panels, CE and TAC strategies perform increasingly poorly, while the POMDP continues to maintain the average stock close to $B_{MSY}$.  Notably, CE is significantly more impacted by measurement uncertainty than TAC, with CE averaging even lower biomass than TAC under moderate measurement uncertainty.  These results confirm that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE, but is less robust to environmental error. In contrast, the POMDP approach successfully handles both sources of uncertainty, maintaining higher stocks, near $B_{MSY}$ level, and producing the highest yield.

>>>>

```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,  dev="cairo_pdf", cache = 1)
library(tidyverse)
library(gridExtra)
```

@@

Imperfect information is ubiquitous in ecological management and conservation decision making. While the pressing concerns of global change have put the spotlight on forecasting, i.e. the uncertainty of the future [e.g. @Petchy2015], management decisions must also contend with uncertainty of the present: How many fish are in the sea today? What regions harbor the most fragile biodiversity? How do we make use of a rapidly expanding but opportunistically collected data, such as in citizen science efforts, which may involve strong sampling biases? Much of our focus in dealing with uncertainty has been through the lens of statistical approaches to model fitting.  The rise of approaches such as Hierarchical Bayesian Modeling have given us sophisticated tools to reflect uncertainty in our data and uncertainty in our knowledge of mechanisms and parameters in our models [@Ellison2004].  But estimating a model is not the same as making a decision, particularly when a model leaves us with considerable uncertainty about current and future states. A model can only provide the rules of the chess board.  It is the role of decision theory to provide us with a strategy.

Decision theory has a long history in the ecological literature [e.g @Schaefer1954; @Walters1978; @Clark1990; @Shea1998; @Polasky2011] which has sometimes been overshadowed by recent emphasis on model fitting.  While in principle a model will be updated throughout the decision process (as we will discuss), it is convenient to think of the model as "given," estimated, with uncertainty, using the best available methods. The goal of a decision theory is not maximize some abstract notion of model fit (least squares, likelihood, etc), but a more practical notion of utility (happiness, economic value, or any combination of stakeholder objectives, see @Halpern2013).   Approaches for decision making under uncertainty in ecological systems can be divided into two camps: (1) those based upon "optimal control" solutions, and usually favored by natural resource economists, and (2) those based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011]. While researchers have recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknowledged by heuristic methods, computational barriers to doing so have stymied this progress.  We illustrate how the limitations of these approaches have manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigor of optimization to resolve a long-standing paradox and suggest a more robust approach to management.

@@

In this paper, we examine three strategies to managing a natural population, such as a fishery, under such uncertainty.  This problem has a long history in both theory and practice, where seemingly innoccous simplifying assumptions have led to a deep paradox in how managers should respond to uncertainty.  The precautionary principle [@Kriebel2001] would seem to suggest that the greater the uncertainty in our models, the more cautious we should be in management policies.  Yet a well-known mathematical proof by @Reed1979 established that under quite general conditions, the optimal strategy under uncertainty is no different than under no uncertainty at all.  Attempts to resolve this have since only deepened this paradox, arguing that adding additional sources of uncertainty, such as imperfect measurements, should lead to optimal harvest that increases with greater uncertainty [e.g. @Clark1986; @Sethi2005].  Meanwhile, much of actual practice has (perhaps fortunately, as we shall see) ignored these results in favor of a more heuristic application of the precautionary principle.  Here, we show that this paradox can be resolved by uncovering previous shortcuts and tackling the optimal management problem head-on, which we can solve thanks to the computational efficiency of sophisticated algorithms developed more recently in the field of robotics.  This approach allows us to reject this paradox largely in favor of the precautionary principle. Further, our comparisons show that widely recognized existing approaches in both theory and practice are not by themselves sufficient to manage a population under substantial uncertainty over the long term, but our recently borrowed approach from robotics literature can under identical conditions sustain both higher biomass and higher economic yields than those approaches.  This conclusion has important implications not only for fisheries but eocsystem more broadly, underscoring the danger of not correctly accounting for measurement uncertainty when making decisions, and further to the significant opportunity to learn from fields such as robotics when faced with complex and uncertain decisions.

@@

There is one further subtlety arising in this seemingly simple problem statement that we must address. Given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?  Following convention in the optimal control literature (e.g. [@Reed1979; @Clark1990; @Sethi2005]), we will assume the latter; measurement occurs immediately before harvest, followed by recruitment. This assumption has occasionally been reversed to introduce some additional uncertainty from stochastic recruitment immediately prior to setting the harvest quota [e.g. @Clark1986; @Weitzman2002], which appears to mimic uncertainty in measurements but does not allow uncertainty to accumulate since the measurement step still occurs without error.

@@

## Simulations

We consider the average fish biomass across 100 replicate simulations of the same stock dynamics (Eq 4) under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.  Each suite of simulations is considered under three sequentially higher measurement error regimes, from $\sigma_m = 0$ (no measurement errors), $\sigma_m = 0.1$ (low error), and $\sigma_m = 0.15$ (moderate error), as indicated, while stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.  Empirical estimates of measurement error in stock sizes vary widely, with estimates often dependent upon methodological assumptions.  Early estimates have put possible error as high as 50% [@Clark1986] and more recent analyses have suggested an average of closer to 36%, though with cases that may be much higher [@Ralston2011].  Consequently, range between 0 to 15% provides a probably conservative estimate of typical unceratinty ranges, which we show is nevertheless large enough to drive substantially negative ecological and economic outcomes under management strategies that do not appropriately account for that error over the long run.

@@

In the absence of either environmental noise or measurement error POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level (see Supplemental material, Figure S1).  The introduction of stochastic growth has a significant negative impact on the TAC strategy, with a mean biomass significantly lower than $B_{MSY}$ (first panel, Figure \ref{sims}; this impact is even more severe for MSY, see Figure S2).  Without measurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.  As measurement error increases in the subsequent panels, CE and TAC strategies perform increasingly poorly, while the POMDP continues to maintain the average stock close to $B_{MSY}$.  Notably, CE is significantly more impacted by measurement uncertainty than TAC, with CE averaging even lower biomass than TAC under moderate measurement uncertainty.  These results confirm that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE, but is less robust to environmental error. In contrast, the POMDP approach successfully handles both sources of uncertainty, maintaining higher stocks, near $B_{MSY}$ level, and producing the highest yield.
